Explaining Black-Box Models through Counterfactuals

### 2023-11-16 examples/CounterfactualExplanations.jl/paper.qmd

---
title: "Explaining Black-Box Models through Counterfactuals"
keywords:
  - Julia
  - Explainable Artificial Intelligence
  - Counterfactual Explanations
  - Algorithmic Recourse
authors:
  - name: Patrick Altmeyer
    orcid: 0000-0003-4726-8613
    affiliation: Delft University of Technology
  - name: Arie van Deursen
    affiliation: Delft University of Technology
  - name: Cynthia C. S. Liem
    affiliation: Delft University of Technology
abstract: |
    We present `CounterfactualExplanations.jl`: a package for generating Counterfactual Explanations (CE) and Algorithmic Recourse (AR) for black-box models in Julia. CE explain how inputs into a model need to change to yield specific model predictions. Explanations that involve realistic and actionable changes can be used to provide AR: a set of proposed actions for individuals to change an undesirable outcome for the better. In this article, we discuss the usefulness of CE for Explainable Artificial Intelligence and demonstrate the functionality of our package. The package is straightforward to use and designed with a focus on customization and extensibility. We envision it to one day be the go-to place for explaining arbitrary predictive models in Julia through a diverse suite of counterfactual generators.
format:
  juliacon-proceedings-pdf: default
bibliography: ref.bib
execute:
  eval: false
  echo: false
  output: false
jupyter: julia-1.9
---

>>>>

---
title: "Explaining Black-Box Models through Counterfactuals"
keywords:
  - Julia
  - Explainable Artificial Intelligence
  - Counterfactual Explanations
  - Algorithmic Recourse
authors:
  - name: Patrick Altmeyer
    orcid: 0000-0003-4726-8613
    affiliation: Delft University of Technology
  - name: Arie van Deursen
    affiliation: Delft University of Technology
  - name: Cynthia C. S. Liem
    affiliation: Delft University of Technology
abstract: |
    We present `CounterfactualExplanations.jl`: a package for generating Counterfactual Explanations (CE) and Algorithmic Recourse (AR) for black-box models in Julia. CE explain how inputs into a model need to change to yield specific model predictions. Explanations that involve realistic and actionable changes can be used to provide AR: a set of proposed actions for individuals to change an undesirable outcome for the better. In this article, we discuss the usefulness of CE for Explainable Artificial Intelligence and demonstrate the functionality of our package. The package is straightforward to use and designed with a focus on customization and extensibility. We envision it to one day be the go-to place for explaining arbitrary predictive models in Julia through a diverse suite of counterfactual generators.
format:
  juliacon-proceedings-pdf:
    keep-tex: true  
  juliacon-proceedings-html: default
bibliography: ref.bib
execute:
  eval: false
  echo: false
  output: false
jupyter: julia-1.9
---
